---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r}
options(digits = 2)
set.seed(1234)
```

```{r}
library("arulesViz")
data("Groceries")
str(Groceries)
Groceries@itemInfo #this gives the item catergories 
```
##Frequent items

```{r}
summary(Groceries@itemInfo$level1) #gives the number of named grocery items are in a specific category

unique(Groceries@itemInfo$level2) #uniquely identifies the categories in the data
```
```{r}
Groceries@itemInfo
```


```{r}
# the code below plots a frequency plot on grocery items
itemFrequencyPlot(Groceries,topN=10, col=heat.colors(10,alpha=0.8),horiz=T,type="absolute",cex.names = 1.2,xlab = list("Item frequency", cex=1.2))

title(main=list("Most frequent items",cex=1.5, font=4)) 
```

##most frequented basket 


```{r}
frequentItems <- eclat (Groceries, parameter = list(supp = 0.01,minlen=3)) #itemsets that appear one percent of the time (at least 98 times of size 3 items and above)

summary(frequentItems) #(this it to get the most common basket size)

freqItemsSupp<-inspect(head(sort(frequentItems,by="support", decreasing = T))) 
```


```{r}
frequentItemsSize3 <- eclat (Groceries, parameter = list(supp = 0.01, minlen=3 , maxlen=3 )) #find the most common basket with size 3 items (since the top 5 basket sizes only had 3 items)

frequentItemsSize3 

freqItemsSupp<-inspect(head(sort(frequentItemsSize3 ,by="support", decreasing = T)))
```



## examine rules with certain criteria

```{r}
rules <- apriori(Groceries, parameter=list(support=0.001,conf=0.8, minlen=2, maxlen=16)) #frequency of occurrence at least 1 in 1000, and with 80% confidence - the combination of items should be found together at least 80% of the time 

surules <- subset(rules,subset=items %pin% c("whole milk","other vegetables","soda", "yogurt", "rolls/buns")) #frequency of occurrence at least 1 in 1000 and will be used for the grouped data graph - these rules should include the mentioned grocery items 

#inspect(surules)

rules1 <- apriori(Groceries, parameter=list(support=0.001, minlen=5)) #output itemsets that appear atleast 1 percent of the time and at least have 5 items 

```


```{r}
summary(rules)

summary(rules1)
```

```{r}
inspect(head(rules, n = 4, by =c("confidence","lift"))) #high confidence and lift to show the most probable item combinations


inspect(head(rules1, n=3,by ="confidence"))
```

```{r}
inspect(head(rules, n = 10, by ="support")) #show the top 10 rules that satisfied the previous requirements and sort by their support (probability of finding the combination)
```
```{r}
inspect(head(rules, n = 3, by ="confidence"))
```



##Scatterplot
```{r}
plot(rules)
```

```{r}
head(quality(rules))
plot(rules, measure = c("support", "lift"), shading = "confidence")
# #Figure 2 shows this plot with lift on the y-axis. Here it is easy to identify all rules with high
# lift.

```

Here support and confidence are used for the x and y-axes and the color of the
points is used to indicate order," i.e., the number of items contained in the rule.
```{r}
plot(rules, method = "two-key plot") 
# #From the plot it is clear that order and
# support have a very strong inverse relationship, which is a known fact for association rules
```



##  matrix-based method 
```{r}
subrules <- rules[quality(rules)$confidence > 0.8]

subrules #371 rules that satisfy the above requirement

plot(subrules, method = "matrix", measure = "lift") #thus about 371 rules represented 

#the vertical axis is the consequent corresponding to certain items (6 itemsets)

#the horizontal axis corresponds to each subrule's antecedent 

#basically almost all have whole milk in their consequence 


```

With a method like this we could see which consequent item is the most found among from the most popular rules, although somewhat hard to see, so I suggest playing around with the code here

```{r}
plot(subrules, method = "matrix3D", measure = "lift")
```


##grouped matrix-based methods
```{r}
plot(surules, main="",horiz=T, method = "grouped",gp_labels = gpar(col = "red", cex=0.8, fontface="italic"), control = list(k = 10,rhs_max=3,col=heat.colors(4))) #can extract it better in terminal  - look for line 63 that specifies what surules is 

png(file = "myplot2.png",  width = 600, height = 700, bg = "white",res=96)
 plot(surules, main="", legend=" ",gp_labs	 =  list(cex = 1, fontface = "bold", font = c(bold = 2)),  method = "grouped",gp_labels = gpar(col = "grey1", cex=0.7, fontface="italic"), control = list(k = 5,rhs_max=10,col=heat.colors(1)))
dev.off()
```


I found this the most fascinating and it shows left hand side(precedent) vs right hand side (consequent). The size of the circle corresponds to the popularity of this combination ( the bigger the more frequent). I suggest playing around with colours here as well. Function would only plot properly in terminal.
```{r}
plot(rules, method = "grouped", control = list(k = 50), cex=0.5)
```


##graph-based method 
This method also has a pretty cool visualisation where we can see which item relates to what, where the arrow points to the rhs item and the random circles inbetween indicate size of lift. 
```{r}

require("arulesViz")
subrules2 <- head(rules, n = 5, by = c("confidence","lift")) #top ten items using lift

plot(subrules2, method = "graph",control=list(type="items",main=" ", nodeCol="yellow", edgeCol="blue", engine = "interactive", shading = NA))

plot(subrules2, method = "graph",control=list(nodeCol="yellow", edgeCol="blue", type="items", layout=igraph::with_kk()))

```

```{r}

##various variations in the setup of the graph
plot(subrules2, method="graph", layout=igraph::in_circle())
plot(subrules2, method="graph", layout=igraph::with_kk())
plot(subrules2, method="graph", layout=igraph::with_fr())
plot(subrules2, method="graph", layout=igraph::with_graphopt())
plot(subrules2, method="graph", layout=igraph::with_gem())
plot(subrules2, method="graph", layout=igraph::on_grid())
plot(subrules2, method="graph", layout=igraph::on_sphere())
plot(subrules2, method="graph",
layout=igraph::with_graphopt(spring.const=5, mass=50))

#nodeCol = grey.colors(10), edgeCol = , alpha = 1
plot(subrules2, method="graph", control=list(nodeCol=heat.colors(1), edgeCol="orange2",layout=igraph::with_dh(maxiter=1250)))
title(main=list("10 Most probable item combinations",cex=1.5, font=4))
```


```{r}
png(file = "myplot.png",  width = 600, height = 480, bg = "white",res=80)
plot(subrules2,main=" ",method="graph",shading = NA, control=list(nodeCol=heat.colors(1), edgeCol="orange2",layout=igraph::with_dh(maxiter=1250),cex=1.2)) #printing of the graph plot
dev.off()
```

```{r}
png(file = "myplot2.png",  width = 600, height = 480, bg = "white",res=80)
plot(subrules2,main=" ",method="graph",shading = NA, control=list(nodeCol=heat.colors(1), edgeCol="orange2",layout=igraph::in_circle(),cex=1.2))
dev.off()
```


```{r}

subrules2 <- head(rules, n = 10, by = "lift") #top ten items using lift

plot(subrules2, method = "graph", engine='interactive', shading = NA) 

```



##coordinate system 

The width of the arrows represents support and the intensity of the color represent confidence. It is obvious that for larger rule sets visual analysis becomes difficult since with an increasing number of rules also the number of crossovers between the lines increases.
```{r}
plot(rules, method = "paracoord")
```

Interesting visualisation that could be difficult to intepret  
```{r}
plot(subrules2, method = "paracoord", control = list(reorder = TRUE)) #reorder the set such that there are less crossovers

```

## double-decker methods 

```{r}

rules <- apriori(Groceries, parameter=list(support=0.001, minlen=3, maxlen=5))

oneRule <-head(subset(rules, subset=items %ain% c("whole milk","other vegetables","root vegetables")), n=1, by=c("confidence","lift"))


inspect(oneRule)

png(file = "myplot2.png",  width = 600, height = 480, bg = "white",res=80)
plot(oneRule, main=" ", method = "doubledecker", data = Groceries) #at the bottom of graph we have the antecedent items in the rule while on the right side we have the consequent items
dev.off()

```

This could be used for a specifc rule where one wants to see a specific pattern or combination of antecdent and consequent 


